---
title: "class15_markdown"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

Install DESeq2 from bioconductor
Say no when asks if want to install from source or update
```{r}
## Copy paste into console one at a time
##source("http://bioconductor.org/biocLite.R")
##biocLite()

# For this class, you'll also need DESeq2:
##biocLite("DESeq2")
```
Run from library
```{r}
library(BiocInstaller)
```

Read the csv files count and metadata files
```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

understand the data by looking at the heads
```{r}
# head(counts)
# head(metadata)
# dim(counts)
# dim(metadata)
```
matches except that have ensemble gene name in first column

want to check if columns match rows
```{r}
colnames(counts)
metadata$id
colnames(counts)[-1] == metadata$id
#use all to check all cases and return single function
all(colnames(counts)[-1] == metadata$id)
```

First task is find which columns in counts correstponds to control conditions (i.e. no drug).
Then calculate average count value for each gene.
```{r}
control.inds <- metadata$dex == "control"
# use ID to look up columns and use rownames to match
control.ids <- metadata$id[control.inds]
head(counts[,control.ids])
```

Calculate average of each row
```{r}
control.mean <- rowSums(counts[,control.ids]) / ncol(counts[,control.ids])
names(control.mean) <- counts$ensgene
```

Now do same for drug treated columns
```{r}
treated.inds <- metadata$dex == "treated"
treated.ids  <- metadata$id[treated.inds]
treated.mean <- rowSums(counts[,treated.ids]) / ncol(counts[,treated.ids])
names(treated.mean) <- counts$ensgene

```

Combine meancount to make it cleaner
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

Sum instances
```{r}
colSums(meancounts)
```

Plot treated vs control. If no RMSE then perfect correlation, if scatter then other factors may be related
```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
```

Because a bunch of values very small, better to show as a logarithm plot
```{r}
plot(log(meancounts$control.mean), log(meancounts$treated.mean))
# or can do
# plot(meancounts$control.mean, meancounts$treated.mean, log = "xy")
```

Log2 is usually used because it has nice properties
Creates another column in meancounts that has log comparitive properties
0 is on line, + is above, - is below (above means gene expression goes up)
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

Nan = log(x/0)
-inf = log(0)
want to filter out these odd values, if zeroes then no data on gene
arr.ind retuns array row and column of instance
unique filters out duplicates 
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)
to.remove <- unique(zero.vals[,1])
mycounts <- meancounts[-to.remove,]
```

Create common threshold to determine whether data "different." Using -2 and 2
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

Use sum to get amount of up and down
```{r}
sum(up.ind)
sum(down.ind)
```

There are `sum(up.ind)` up regulated genes and `sum(down.ind)` down regualated genes.

Show up regulated genes
```{r}
head(mycounts[up.ind,])
```
Problem is do not what know proteins are being mentioned

Use merge function to combine annotation data with regulated data
use head to get idea of annotation.data
```{r}
annotation.data <- read.csv("data/annotables_grch38.csv")
# head(annotation.data)
mycounts.annotated <- merge(mycounts, annotation.data, by.x = "row.names", by.y = "ensgene")
# head(mycounts.annotated)
```

Save time and work efficiently by using the DESeq2 analysis
```{r}
library(DESeq2)
citation("DESeq2")
```

Need to populate w count data, metadata, and formula for experiment design
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

Want to run function
will but but then need to get results and data
```{r}
dds.return <- DESeq(dds)
dds.results <- results(dds.return)
summary(dds.results)
```

Volcano plot: significantly regulated gense vs. fold change
shows fold change (x-axis) vs. p-value

Order results by smallest p-value
```{r}
resOrdered <- dds.results[order(dds.results$pvalue),]
```

Change p-value cutoff to 0.05 instead of default 0.1
```{r}
res05 <- results(dds.return, alpha=0.05)
res01 <- results(dds.return, alpha=0.01)
summary(res05)
```

More generic way of accessing acutal subset of data framce
```{r}
resSig05 <- subset(as.data.frame(dds.results), padj < 0.05)
nrow(resSig05)
resSig01 <- subset(as.data.frame(dds.results), padj < 0.01)
nrow(resSig01)
```

Arrange and view results by adjusted p-value
```{r}
ord <- order( resSig01$padj )
#View(res01[ord,])
head(resSig01[ord,])
```

Write out ordered significant results with annotations as csv
```{r}
write.csv(resSig05[ord,], "data/signif05_results.csv")
write.csv(resSig01[ord,], "data/signif01_results.csv")
```

**Data Visualization**
Plotting
```{r}
mycols <- rep("gray", nrow(res01))
mycols[ abs(res01$log2FoldChange) > 2] <- "red"
inds <- (res01$Spadj < 0.01) & (abs(res01$log2FoldChange) > 2)
mycols[inds] <- "blue"

# Volcano Plot
plot(res01$log2FoldChange, -log(res01$padj), col=mycols, ylab= "-Log(P-value)", xlab= "Log2(FoldChange")

# INC


```


