---
title: "class11_markdown"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Section 1. The PDB
```{r}
stats <- read.csv(file = "data/Data Export Summary.csv")
```

Q1:  Determine the percentage of structures solved by X-Ray and Electron Microscopy. From
the website what proportion of structures are protein?

 
```{r}
percent.by.method <- stats$Total/sum(stats$Total) *100
names(percent.by.method) <- rownames(stats)
```

```{r}
sum(stats$Proteins)/sum(stats$Total) *100
```

Q2: There are 1138 structures of HIV-1 protease in the current PDB

Q3: Only see one molecule becasuse the program groups the molecules together as one


## Section 3: Introduction to Bio3D in R

Load package
```{r}
library(bio3d)
```
Now can use functions in package

```{r}
pdb <- read.pdb("data/1hsg.pdb")
pdb
```

```{r}
pdb$atom
```

Q8 trying to select out only protein and drug only parts of files/molecules

```{r}
pdb <- read.pdb("1hsg")
prot.inds <- atom.select(pdb, "protein")
prot.inds
```
```{r}
pdb$atom[prot.inds$atom,]
```
```{r}
prot.pdb <- trim.pdb(pdb, prot.inds)
write.pdb(prot.pdb, file="protein.pdb")
```

Same thing for drug now
```{r}
lig.inds <- atom.select(pdb, "ligand")
lig.pdb <- trim.pdb(pdb, lig.inds)
write.pdb(lig.pdb, file = "ligand.pdb")
```

