---
title: "class13_markdown"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Load bio3d library
```{r}
library(bio3d)
```

Store 1hsg as file.name
```{r}
file.name <- get.pdb("1hsg")
```
Read pdb file and store as hiv
```{r}
hiv <- read.pdb(file.name)
hiv
```
Question 1: The two non-protein resid values are HOH (127), MK1 (1). Resid corresponds to the amino acids and other things (water and ligand molecules) in the protein structure. The full listing in the protein sequence below.

Isolate protein and ligand and create files
```{r}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, file="1hsg_ligand.pdb")
```
 
Question 2:
 In ADT (Auto Dock Tools) a little hard to see docking site so add H atoms, H is usually omitted because 
 
 Question 3: The charges make sense based upon biochemical amino acid interactions
 
 Processing the all.pdbqt to a PDB format file that can be loaded into VMD
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

Question 4:
The docks look in the right region and seem the fit well, however there are many different configurations. The first file configuration looks very clean.

Calculate the RMSD between docking results and crystal strucutre
```{r}
# res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori, res)
```
Question 5: They look good except only the first configuration of crystal binding is within 1A RMSD for all atoms.

Question 6: To determine the RMSD for heavy atoms only, I would create new variables that are trimmed to only contain non H atoms. Then I would run the RMSD.
```{r}
res.noh <- trim.pdb(res, "noh")
ori.noh <- trim.pdb(ori, "noh")
rmsd(ori.noh,res.noh)
```

EXTRA: Normal Mode Analysis (NMA)

Loading hen egg white lysosome structure and plotting
```{r}
##library(bio3d)
pdb <- read.pdb("1HEL")
modes <- nma(pdb)
plot(modes, sse=pdb)
```

Preparing file to show structural variations in VMD
```{r}
# Visualize NMA results
mktrj(modes, mode=7, file="nma_7.pdb")
```

Question 7: Want to look at most flexible portions of HIV-1 protease
```{r}
##already read pdb of 1hsg and saved as variable "hiv"
hiv.modes <- nma(hiv)
##don't need next line as it is just the plot
##plot(hiv.modes, sse=hiv)
mktrj(hiv.modes, mode=7, file="hiv_7.pdb")
```
Loading in a playing the hiv_7.pdb file in VMD shows that the prtotein is very flexible in the region of the ligand This may play a role in docking calculations as the large flexibility makes it such that many more arragnements off amino acids would need to be considered when looking for the right alignment to dock.
